Whose voices are heard?

### 2024-02-29 paper/title_page.qmd
re-submitted to IJPP

format:
  aft-pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
          \usepackage{setspace}
          \linespread{2}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
bibliography: references.bib
execute:
  echo: false
  eval: true
---

```{r setup}
#| eval: true
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")
library(tidyverse); theme_set(theme_minimal())
library(kableExtra)
library(texreg)
library(margins)
library(grid)
```

<!-- # Introduction -->

Protest is a fundamental democratic tool for citizens to voice dissent about issues in society.
In modern democracies, however, citizens usually only hear about protest through mass media.
And this hearing consists not only of the audience becoming aware of a protest event, but ideally should also involve them receiving the information to *understand* the underlying issue and the causes for the public dissent.
However, most of the literature concerning protest coverage in news media points to a phenomenon referred to as *protest paradigm* to explain a recurring patterns in news about demonstrations:
the media focus on crime, clashes with police and the spectacle surrounding a protest or the protesters themselves and their looks, demographics or occupation, while neglecting to explain *why* protesters spent time and resources on organising and participating in a street demonstration in the first place.
The *protest paradigm* explains this by presuming a mental catalogue, supposedly common among journalists, which indicates to them "where to look (and where not to look), and informs them about what to discover" [@chanJournalisticParadigmsCivil1984, p. 187].
<!-- The *protest paradigm* is thus closely related to the literature on the sociology of newsmaking, which observed that news workers are "routinizing the unexpected" [@tuchmanMakingNewsDoing1973, p. 110] by creating practical and effective procedural rules, routines and mental catalogues of news story themes to cope with the onslaught of new events. -->

Most of the focus in previous studies has been spent on the dangers of this routinised reporting.
This is due to the normative core of the *protest paradigm*:
by regularly focusing on the action itself, rather than the sometimes more elusive and complicated message [@smithProtestAgendaBuilding2001] reporting shaped by the *protest paradigm* could lead to delegitimisation and marginalisation of protest events, protest movements, or protest as a political resource [@mcleodSocialControlSocial1999].
*Some* evidence for the *protest paradigm* has been found in most studies that analysed news coverage of protests over the years [for a systematic overview see @gruberTrouble2023].
However, more recent work has also asked *how common* delegitimising devices of the *protest paradigm* are, as nobody can deny that the press at least occasionally discusses the socio-political issues protesters try to rise awareness for --- hence *legitimising their actions*.
While it seems fairly clear that *some* coverage follows routinized dismissive patterns, this might not be that damaging to protests as long as substantive information about causes and grievances are still available elsewhere.
Studies following this logic suggest that the delegitimising aspects of the *protest paradigm* might occur in reporting more conditionally [@kilgoHierarchy2019; @woutersReportingDemonstrationsEpisodic2015], and less frequently over time [@gruberTrouble2023; @araizaHandsDontShoot2016; @kyriakidouIndignadosProtestsSpanish2017].

The current study follows the footsteps of this line of research by asking:

> **RQ: Which factors explain the choice of frames by the news media when covering protest events?**

It contributes to the literature in two ways.
Firstly, while this is not the first study to test factors which condition protest coverage [e.g., @boyleAdherenceProtestParadigm2012; @kilgoHierarchy2019; @leeTriggeringProtestParadigm2014], the set of variables presented here is novel and, to my knowledge, the most comprehensive so far.
The first section of the article discusses which mechanisms and protest features scholars in the *protest paradigm* tradition have hypothesised to drive coverage and distills them into seven event-, outlet- and time-based factors.

Secondly, while previous studies often followed a dichotomy where coverage is divided into reporting that is either legitimising or delegitimising a protest action, this article tries to disentangle effects on the usage of different frames in reporting.
The main frames of protest reporting are identified and measured in a dataset that spans newspaper coverage of all protest events in the United Kingdom between 1992 and the end of 2017 as identified in the *Mass Mobilization Project*'s [@ClarkMMPdata2019] (*N = 27,496 newspaper articles about 471 events*).
Using a set of multilevel logistic regression models, the article proceeds to test which of the factors from the literature explain why some frames are used more (or less) often in the analysed reporting.
This is discussed for seven individual frames, as identified in the material, and for two grouped dependent variables which indicate the presence of absence of any *legitimising* or *delegitimising* frames.
Grouping frames into this dichotomy allows to link results back to the reviewed literature.

Results indicate that the way protests are framed in news coverage is influenced by four primary factors:
**Violence**: incidences of violence in protests lead to more coverage that delegitimizes the protest and less that legitimizes it.
**Objectives of the protest**: the goals behind a protest play a role in shaping its media portrayal. The connection between these objectives and the framing is intricate, however, with protest goals having a more pronounced effect on legitimizing frames.
**Characteristics of the media outlet**: The nature and ideological leanings of a news organization also affect how protests are covered, although this influence is not as strong as initially thought.
**Timing of coverage**: reports that are more immediate after an event, and reports that are from more recent years tend to include more legitimizing frames in their coverage.

# What Drives Protest Coverage in News Media? {#sec-factors}

To structure the review of previously identified factors that shape protest reporting, I interrogate factors on the event-level, outlet-level and time-related factors below.^[Similar to how @shoemakerMediatingMessage21st2014 divide influences believed to shape the news into several levels ordered from micro to macro.]
It is important to recognize, however, that the factors outlined below are rarely singular determinants of news coverage. Instead, a complex interplay of influences guides journalists in emphasizing specific aspects of a protest event over others. Some of these influences are consciously considered by journalists, while others remain implicit, as demonstrated by @tenenboimweinblattProducingProtestNews2014.

## Event-Level Factors

### Protest Goals

On the level of the individual event, several studies document that the *goals or causes* voiced by protesters shape the coverage these events receive.
In earlier studies, this has been mostly attributed to the idea that the media are agents of social control, who consciously and unconsciously reward conformity to and punish deviance from a perceived mainstream or *status quo* [@mcleodSocialControlSocial1999].
In this regard, @shoemakerMediaTreatmentDeviant1984 found that groups that are perceived as more deviant or radical tend to receive more delegitimising coverage [@hallPolicingCrisisMugging1978]^[Also see the deviance sphere in @hallinUncensoredWarMedia1986.].
Based on this, @boyleInfluenceLevelDeviance2004 tested how the *level of deviance* of protests shape their coverage.
They conclude that coverage of deviant protest conformed closer to the *protest paradigm*, conceptualised as more critical coverage, coverage employing episodic framing and coverage that is less likely to use protesters as sources.
<!-- The idea of a *level of deviance* can also be linked to @bennettindexing1990's theory of "indexing", which postulates that the range of voices and interpretations of events are organised "according to the range of views expressed in mainstream government debate about a given topic" (p. 106). -->
<!-- This means that in times of elite consensus, media will limit its discussion to the opinions expressed by the powerful. -->
<!-- Only in times of elite dissensus, the voices of the broader public will be heard. -->
<!-- Otherwise, alternative information or narratives are assumed to encounter fierce resistance or outright blockage [@bennettNoneDareCall2006]. -->
<!-- Generally, it is thought that indexing hurts protesters' ability to explain the causes of their efforts [@andrewsMakingNewsMovement2010]. -->
However, while the notion that deviant goals are less likely to be cast in a positive light is well supported, newer studies have departed from it due to a theoretical ambiguity:
there is no irrefutable conceptualisation of *the status quo* as political and social realities are constantly changing [@kilgoHierarchy2019; @harlowNewParadigm2023].
A goal that was a threat to the status quo five or ten years ago could be in favour of the current situation.

Analysing the deviance of protest goals from the status quo, therefore, appears less meaningful and has shifted the attention to more clear cut categories.
@boyleInfluenceLevelDeviance2004 and @boyleNewspapersProtestExamination2005 studied a range of events between 1960 and 1999 and found that anti-war protesters were treated most critically, followed by labour protests (e.g., strikes and pickets), protests against wrongdoings by the police [only @boyleInfluenceLevelDeviance2004] and social-issue protests (e.g., abortion or hunger) which were treated most favourably.
In an analysis of newer coverage (2007-2009), @boyleAdherenceProtestParadigm2012 found that *political protests*, by which they mean protests seeking political change, were treated more critically than anti-war and social-issue protests.

Instead of ranking these types of protest according to the level of deviance, @kilgoHierarchy2019 suggest that "journalistic routines are subject to a *hierarchy of social struggle*, in which certain topics are given precedence and legitimacy, and others are delegitimised, trivialised, or ignored altogether" (p. 16).
They also add several new types to this list.
Specifically, they found that coverage of protest against anti-Black and anti-Indigenous racism, international protest, and anti-Trump protests use frames that delegitimise their issue far more often than other protest types.
For some issues, like health, the environment and immigration protests, the press used legitimising frames more often.
Similarly, @harlowWholeWorldWatching2020 tested the relation of different protest categories to the usage of several delegitimising and legitisimising frames.
They find that while not all protest types had a significant influence on the use of frames, some patterns emerged:
conservative protests, for example, were less likely to be covered in a riot frame but more likely to be covered in a spectacle frame, while socio-economic protests were more likely to be covered using the riot frame.
For a legitimising frame, dubbed debate frame, no significant relationship to a protest type was found.

Following the clarifications of the protest goal variable by @kilgoHierarchy2019 and @harlowWholeWorldWatching2020, I expect that some goals are more likely to receive legitimising and some are more likely to receive delegitimising coverage, i.e. that a *hierarchy of social struggle* will emerge in UK reporting of protest.
I will use five broad categories for protest goals, ordered from most to least likely to receive critical coverage these are:
*political*, *anti-war*, *labour*, *police* and *social-issue* protests [@boyleAdherenceProtestParadigm2012; @boyleNewspapersProtestExamination2005; @boyleInfluenceLevelDeviance2004].
The hypotheses I will test are thus:

- `r hyp("H1a")`
- `r hyp("H1b")`

Note that the expected hierarchy in **H1b** is the flipped version of **H1a** as it is generally expected that the likelihood for a protest to be covered with a legitimising frame should be roughly diametrically opposed to the likelihood for a protest to be covered with a delegitimising frame.

### Tactics

The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980].
@boyleMeasuringLevelDeviance2009 showed that when both protest goals and tactics are taken into account, violent or disruptive tactics facilitate critical coverage stronger than protest goals.
<!-- This insight had already been observed by one of the earliest studies of protest coverage: -->
<!-- @gitlinWholeWorldWatching1980 found that the amount of outright negative coverage, instead of trivialisation, increased as anti-war protests became more extreme. -->
@boyleAdherenceProtestParadigm2012, who additionally test the effects of protest location, protest type, and a group's goals in an international comparison of protest coverage, conclude that the tactics employed during a protest are the strongest determinant of delegitimising coverage.
Likewise, @WasowSeeding2020 concludes that nonviolent black-led protests during the civil rights movement led to more coverage of civil rights issues, while protester violence invoked language associated with disorder and calls for law and order.

How extreme the tactics are is usually determined by two variables:
do the protesters break laws and do they engage in violence?
From a broader perspective, a focus on extreme protest tactics is, therefore, in line with the professional norms of "newsworthiness":
usually, the clashes rather than the message of a protest will produce most conflict, drama, unusualness and negativity.
This leads to the following expectations:

- `r hyp("H2a")`
- `r hyp("H2b")`

<!-- The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980]. -->
<!-- When tactics become more violent, coverage becomes less supportive of the cause and more supportive of repressive measures. -->
<!-- Nevertheless, protesters often use violent tactics because this will usually lead to (more) coverage [@boyleAdherenceProtestParadigm2012]. -->
<!-- **H2a** and **H2b**, therefore, also speak to the question if this bargain makes sense for a movement. -->

### State Response

As mentioned above, @WasowSeeding2020 found in his study of black-led protests between 1960 and 1972 that protester-initiated violence led to delegitimising coverage and subsequently turned public opinion and voter preferences against the civil rights movement.
However, @WasowSeeding2020 also tested the effects of violent state response towards the movement.
Specifically, he found that there are four scenarios that yielded slightly different outcomes:
When the state tolerated peaceful protest, this lead to coverage which legitimised the concerns voiced by protesters;
when the state used violent repression against peaceful protesters, this lead to coverage that is more supportive of the protesters' goal --- and to an overall higher volume of coverage about the respective event;
when only the protesters were violent, the civil rights movement was penalised in the coverage;
the same was, however, true when both police and protesters engaged in substantial violence.
In this last case, the coverage is predicted to focus on concerns about order and public safety.

Wasow's [-@WasowSeeding2020] study thus mostly confirms the relationship between violence and (de)legitimising coverage formalised in **H2a** and **H2b**.
A scenario in which the state uses violent repression against peaceful protesters, however, has not been covered by any previous studies.
Following @WasowSeeding2020, I include this additional layer, which leads to the expectation that:

- `r hyp("H3")`

## Outlet-Level Factors

### Newspaper Ideology

In the landmark study which defined the idea of a *protest paradigm*, @chanJournalisticParadigmsCivil1984 found that not all newspapers engage in delegitimising protests in the same way:
the left-leaning media were less hostile towards the protesting teachers and students in Hong Kong than centrist media, which were in turn less hostile than right-leaning journalists.
<!-- This factor has usually been overlooked in subsequent US-based studies, which might have assumed that ideological differences between major outlets are less important given the emphasis on objectivity in the country's journalistic culture. -->
More recent studies, however, have started again to employ newspaper ideology as an explaining factor for the level of delegitimising reporting different protests face [@leeTriggeringProtestParadigm2014; @shahinProtestingParadigmComparative2016; @ozMainstreamMediasCoverage2016].
UK media --- the focus of this study --- are known to be distinctly more partisan than outlets in the US and elsewhere [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004], which could increase the effect of newspaper ideology on protest reporting.
This leads to the expectation:

- `r hyp("H4a")`
- `r hyp("H4b")`

@weaverRevisitingProtestParadigm2012 add nuance to this:
more than by their right or left ideology, the coverage of an outlet might be determined by the distance of their own ideological tilt to the ideology of a protest in question.
They test this using prime-time cable news coverage about the right-leaning *Tea Party* movement in the United States in 2009-2010 to assess which outlets used marginalisation devices more often.
Their finding suggests that liberally aligned MSNBC was most likely to marginalise the *Tea Party*, followed by the centrist AP and CNN, and Fox News being supportive overall.
Since protests have traditionally held goals that were considered left more often in Western countries --- and continue to do so [@borbathDifferentWorldsContention2020] --- the assumption that right-leaning media are generally more critical of protest might be incorrect.
This adds two hypotheses:

- `r hyp("H5a")`
- `r hyp("H5b")`

### Type of Newspaper

Gavin [-@gavinPressTelevisionBritish2007, pp. 95-118] and @rosieAnarchistsWorldCup2009 add another factor related to newspaper outlets:
the difference between tabloid and broadsheet newspapers.
In general, it is expected that tabloid newspapers are more focused on human interest and entertainment [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004].
And indeed, @gavinPressTelevisionBritish2007, who assessed the May Day demonstration in London in 2000, concluded that this focus translated to tabloid outlets using the delegitimising elements from the *protest paradigm* more extensively and failing to mention any motivation or goals of the protest.
He also found that broadsheet newspapers --- especially *The Guardian* --- made an effort to explain why the protests were held.
In their qualitative study, @rosieAnarchistsWorldCup2009 find that tabloid newspapers were more extreme in their framing of peaceful G8 protesters as a fringe group of radical militants.
<!-- The sharp divide in reporting style between broadsheets newspapers (*The Daily Telegraph*, the *Financial Times*, *The Guardian*, *The Independent* and *The Times*), and tabloids (*Daily Express*, *Daily Mail*, the *Daily Mirror* and *The Sun*) is a well-known factor in the UK media system and should hence not be neglected. -->
The sharp divide in reporting style between broadsheets newspapers and tabloids is a well-known factor in the UK media system and should hence not be neglected.

Following the work by @gavinPressTelevisionBritish2007 and @rosieAnarchistsWorldCup2009, the following expectations can be made:

- `r hyp("H6a")`
- `r hyp("H6b")`

## Time-Bound Factors
### Year of Protest

Two factors related to time are expected to have an influence on reporting about protests.
The first one is that protests that happened more recently receive more legitimising coverage.
In an earlier study, I showed that the level of legitimising framing in news about domestic protest has been steadily growing since the early 1990s [@gruberTrouble2023].
Several mechanisms have been suggested for this:
protest has largely been normalised since the first seminal studies about movements in the 1960s have been conducted [@daltonDemocraticChallengesDemocratic2004; @sandersDynamicsProtestBritain2003; @norrisDemocraticPhoenixReinventing2002, pp. 188-212; @vanaelstWhoThatWo2001; @oliverPoliticalProcessesLocal2000].
Today, many different groups employ street protest as a political communication tool, and many more people today have participated in a protest at least once in their lives [e.g., @borbathDifferentWorldsContention2020; @HutterChallenges2019].
Besides this, information about protesters' goals and grievances are more readily available on the internet [@humprechtMoreSameMarketplace2013; @posterWhatMatterInternet2001; @boydSocialNetworkSites2011], while journalists likely feel more pressure to paint a complete picture today as audiences have a sharply increased pool of alternative information sources [@vanaelstPoliticalCommunicationHighchoice2017].

Whether it is a combination of mechanisms or other unknown forces, the expected result of this factor is:

- `r hyp("H7")`

### Days Since Protest

The second factor very likely only affects large scale, highly mediatised protests.
@gottliebProtestNewsFraming2015 suggests that coverage follows a *news framing cycle*:
coverage of protest begins with a focus on conflict --- hence marginalising or ignoring the goals and grievances of a protest --- then shifts to more substantial coverage before returning to conflict.
This concept, based on @downsAttentionCycle1972, follows the logic that journalists will use different frames to cover an event as time passes "to keep the story alive and fresh" [@chyiMediaSalienceProcess2004, p. 22].

The *news framing cycle* intuitively makes sense, yet the proposed time frame might suffer from one problem:
@gottliebProtestNewsFraming2015 suggests that the second phase, when reporting shifts to more substantial coverage, begins in the third week after the start --- that is, at a time when the media will have already forgotten most protest events.

Nevertheless, the idea of a protest news framing cycle seems worth testing, albeit in an abbreviated form:

- `r hyp("H8a")`
- `r hyp("H8b")`

# Data and Methods

To test these hypotheses, I combine two datasets.
The first was created through content analysis and contains measures of framing in reports about a large number of events in UK mainstream newspapers from 1992 to the end of 2017.
The second one contains information about protest events.
These data are matched and the hypotheses are tested through multilevel logistic regression models.
@tbl-variables shows the dependant and independent variables of these models --- one for the presence or absence of each frame.
How this data was sourced is explained below.

```{r variables}
#| label: tbl-variables
#| tbl-cap: Model Variables
variables <- rio::import("../misc/variables.csv")  |>
  mutate(across(everything(), ~gsub('"', '', .x, fixed = TRUE)))  |>
  select(-`Levels-long`, -Mean, -Category)

if (isTRUE(knitr:::pandoc_to() == "latex")) {
  variables  |>
    custom_df_print(caption = "Model Variables",
                    font_size = 9,
                    footnote = "Reference category",
                    latex_options = "scale_down")  |>
    pack_rows("Dependent", 1, 1)  |>
    pack_rows("Event-Level", 2, 4)  |>
    pack_rows("Outlet-Level", 5, 7)  |>
    pack_rows("Time-Bound", 8, 9)  |>
    pack_rows("Level 2", 10, 10)  |>
    column_spec(5, width = "3.7cm")
} else {
  variables  |>
    custom_df_print(caption = "Model Variables", font_size = 9)
}
```

## Content Analysis

I use media coverage of protests as the dependent variable.
Newspaper articles were employed instead of alternatives, such as online news, social media content or TV transcripts for two reasons.
Firstly, despite declining audiences, mainstream news media such as newspapers are still deemed the most important platforms for public debate and the creation of the public agenda [@rogstadTwitterJustRehashing2016].
Newspapers might have lost some of their intermedia agenda-setting power, but the more in-depth writing produced by newspapers nevertheless forms a crucial share of what is redistributed via online news and social media and is key in initiating, amplifying and sustaining attention to an issue [@Langerwindrush2020; @harderIntermediaAgendaSetting2017; @rogstadTwitterJustRehashing2016].
Secondly, suitable data from UK newspapers were already archived and digitally available for the whole period of interest.

@tbl-outlets presents the selection of newspaper outlets, strategically chosen to contain variance on the two important factors:
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing the keywords "protest" and "demonstration" and several misspellings.
<!-- A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles. -->
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, I opted to limit the data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

```{r outlets}
#| label: tbl-outlets
#| tbl-cap: Selection of Newspaper Outlets for Data Collection
outlets <- tibble::tribble(
  ~Type,        ~Ideology,  ~Outlet,                 ~"Available from", ~"Circulation (2005 -- 2017)",
  "Tabloid",    "Left",    "The Mirror",             1995,              "1,720,000 -- 692,295",
  "Tabloid",    "Right",   "The Sun",                2000,              "3,273,000 -- 1,602,320",
  "Tabloid",    "Right",   "Daily Mail",             1992,              "2,426,000 -- 1,442,924",
  "Broadsheet", "Left",    "The Guardian",           1992,              "366,645 -- 153,431",
  "Broadsheet", "Left",    "The Independent",        1992,              "263,595 -- NA",
  "Broadsheet", "Right",   "Financial Times",        1992,              "419,386 -- 190,046",
  "Broadsheet", "Right",   "Daily Telegraph",        1992,              "460,585 -- 907,329",
  "Broadsheet", "Right",   "The Times",              1992,              "679,190 -- 440,736"
)

footnotes = c(
  "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
  "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016."
)
library(gt)
outlets |>
  gt() |>
  tab_footnote(
    footnote = "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
    locations = cells_column_labels(columns = "Circulation (2005 -- 2017)")
  ) |>
  tab_footnote(
    footnote = "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016.",
    locations = cells_body(columns = "Circulation (2005 -- 2017)", rows = 5L)
  )
```

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from Author (Year). <!-- @gruberTrouble2023. -->
In short, frames were identified by coding the frame elements found in Entman's [-@entman1993] definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles.
These elements were then combined to frames using factor analysis.
The result is one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models on the manually coded random sample, all remaining articles were coded using these models.

@fig-frames show the seven identified frames and how often they were coded in news articles.
The majority of them were determined to delegitimise protest using many of the same patterns identified in the *protest paradigm* literature.
They portray protesters as a fringe group of freaks who disturb the social norms (dubbed "decay of morals"), highlight the violence caused by protesters ("troublemakers"), suggest that protest is a bothersome interruption of everyday activities ("nuisance"), or focus on police actions to reinstate public order ("law & order").
Two legitimising frames, on the other hand, describe the main topic and cause of a protest event ("cause & grievances") or highlight the accomplishments of the protesters for a just cause ("righteous struggle").
The remaining frame, "Police Violence", can neither be said to be legitimising nor delegitimising, highlighting the shortcomings of this dichotomy [@harlowNewParadigm2023].
The frame focuses on violence, which is usually assumed to distract from protest demands and hence delegitimising.
Contrary to most theoretical expectations, however, the frame portrays police as responsible for violence.
This could lead to more interest in and sympathy towards a protest goal.

```{r}
#| label: fig-frames
#| fig-cap: Number of Articles in Which Each Frame Is Present
readRDS("../misc/frame_distribution.rds") %>%
  ggplot(aes(y = frame, x = n, fill = super)) +
  geom_col() +
  scale_fill_manual(values = jbg_pal(sel = c("darkred", "darkblue", "black"))) +
  scale_x_continuous(labels = scales::comma) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())
```

<!-- In its most advanced, and probably broadest, conceptualisation, @mcleodSocialControlSocial1999 state that the *protest paradigm* combines *unfavourable story framing*, *reliance on official sources and official definitions* and the *invocation of public opinion* (e.g., the protesters are a minority). -->
<!-- "Protest coverage research needs to move beyond the dichotomous positive versus negative framing to account for the ways in which violence can in fact be a legit- imate way to attract media attention and legitimize protesters’ grievances and tactics (Cammaerts 2013; Drury and Reicher 2018; Saavedra and Drury 2019)" -->

## Independent Variables

This article tests the relation between framing of protest and eight independent variable identified from the literature above:
*Protest Goal*, *Protester Violence*, *State Repression*, *Newspaper Ideology*, *Newspaper Type*, *Ideological Divide*, *Days Since Start*, and *Year of Protest*.
*Newspaper Ideology* were obtained through the meta information of the coded news articles *Newspaper Type*.
For all other variables the *Mass Mobilization Project*'s [@ClarkMMPdata2019] dataset on protest events was employed.
This dataset was chosen for practical reasons as it contains the variables necessary and covers the chosen time frame (1992-2017).
@ClarkMMPdata2019 use the keywords "Protest", "Demonstration", "Riot", and "Mass Mobilization" on the *LexisNexis* newspaper archive, selecting "All News" as the source.
For the selected time frame in the UK, the *Mass Mobilization Project* lists 471 events between 1992 and the end of 2017, as shown in @fig-mmpyear.

```{r}
#| label: fig-mmpyear
#| fig-cap: "Number of Protest Events in the UK per Year per Clark and Regan (2019)"
readRDS("../misc/protest_event_data_uk.Rds")  |>
  mutate(year = lubridate::floor_date(date_start, "year"))  |>
  group_by(year)  |>
  count()  |>
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = jbg_pal(sel = "grey")) +
  scale_x_date(labels = lubridate::year, breaks = "year",
               expand = c(0, 0)) +
  #scale_y_continuous(labels = scales::percent, limits = c(0, 0.7)) +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```

I merged the two datasets by using the time frame of each event and some keywords (see Supplementary Information file).
Overall, 5,639 articles could be matched with the 471 events from the database.
Using the variables listed in @tbl-variables, I then fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectively constitute a panel dataset, since observations must be seen as nested and not independent.
<!-- I used newspaper-year as the Level 2 variable ($n = 180$), since the number of newspapers alone was considered too low for this design. -->
<!-- It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets. -->
<!-- However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016]. -->
<!-- Therefore, I used newspaper-year as the Level 2 variable ($n = 180$). -->
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->
<!-- I assessed whether adding the newspaper-year level increases or decreases model fit. -->
<!-- To compare models, the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) are calculated and compared using Analysis Of Variance (ANOVA) tests. -->
<!-- The lower the AIC and BIC, the better the model fits the data. -->
<!-- Both AIC and BIC are always considerably lower when including the second level into the modelling. -->
<!-- The p-values suggest that including the random effect improves all models significantly. -->
<!-- I examined if adding a newspaper-year level improves model fit. -->
<!-- The model comparisons were done using AIC, BIC, and ANOVA tests. -->
The inclusion of the second level significantly lowers both AIC and BIC for all individual and grouped frame models.
<!-- , and the p-values indicate a significant improvement in all models. -->

# Results

Discussion of results is divided into two steps.
I first refer to the frames aggregated into legitimising and delegitimising to test the hypotheses set up above.
This, however, neglects some of the details revealed through testing the relations to the individual seven frames, which is hence done in the following section.

## (De)Legitimising Framing

```{r}
#| results: asis
#| label: tab-regtwoframes
models <- readRDS("../misc/models.Rds")
model_logit_dele_mlm <- models$model_logit_dele_mlm
model_logit_le_mlm <- models$model_logit_le_mlm

if (is_latex()) {
  texreg(
    c(model_logit_dele_mlm, model_logit_le_mlm),
    caption = "Regression Results for Aggregated Frames",
    custom.model.names = c("Delegitimising Frames", "Legitimising Frames"),
    custom.coef.names = c(
      "(Intercept)",
      "goal: anti-war",
      "goal: labour protests",
      "goal: police",
      "goal: social-issue",
      "violent protest",
      "repression of peaceful p.",
      "right-wing",
      "tabloid newspaper",
      "ideological divide: conflict",
      "ideological divide: ambiguous",
      "days from start",
      "year of protest"
    ),
    groups = list("Event-Level Factors" = 2:7, "Outlet-Level Factors" = 8:11, "Time Bound Factors" = 12:13),
    digits = 3,
    stars = c(0.001, 0.01, 0.05, 0.1),
    symbol = "+",
    #scalebox = 1,
    siunitx = TRUE,
    single.row = FALSE, # True looks better here but much worse on the other. To keep it consistent
    booktabs = TRUE,
    label = "tab-regtwoframes",
    use.packages = FALSE
  )
} else {
  data.frame(placeholder = "In the PDF, the table is around here") %>%
    custom_df_print(caption = "Regression Results for Aggregated Frames")
}
```

```{r}
#| eval: true
model_margins <- readRDS("../misc/model_margins.Rds")
twofram_marg <- list(model_margins$model_logit_dele_mlm,
                     model_margins$model_logit_le_mlm) %>%
  lapply(function(m) {
    summary(m) %>%
      as_tibble() %>%
      filter(p < 0.05) %>%
      mutate(model = colnames(m)[1])
  }) %>%
  bind_rows() %>%
  mutate(factor = clear_var_names(factor))

marg <- function(...) {
  marg_default(df = twofram_marg, ...)
}
```

Table \ref{tab-regtwoframes} shows the models for the legitimising and delegitimising frame aggregates.
Since raw values of coefficient estimates are less useful in a logisitc regression, I additionally provide Average Marginal Effects (AME) in @fig-margtwoframes.
Holding all other covariate values constant, AMEs express here by how much the probability of using a frame increases or decreases *on average*, if a dependent variable changes by 1 in their respective units.
AMEs were calculated using the *margins* package in R [@marginspackage].

```{r}
#| label: fig-margtwoframes
#| fig-cap: Average Marginal Effects for Significant Variables in Aggregated Frame Models
library(ggh4x)
strip <- strip_themed(text_x = list(element_text(colour = "white")), 
                      background_x = elem_list_rect(fill = c(
                        "delegitimising" = "darkred", 
                        "legitimising" = "darkblue"
                      )))

twofram_marg %>%
  ggplot(aes(x = AME, y = factor, xmin = lower, xmax = upper)) +
  geom_point(colour = "#939597") +
  geom_pointrange(colour = "#939597") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "#F65058") +
  facet_wrap2(~model, ncol = 1, scales = "free", strip.position = "top",
              strip = strip) +
  # facet_grid(rows = vars(model), scales = "free_y", space = "free_y")
  labs(y = NULL, x = NULL) #+
  # scale_y_discrete(labels = function(x) str_wrap(x, 15))

# fcsts <- twofram_marg %>%
#   count(model) %>%
#   pull(n)
# 
# margtwoframes <- ggplotGrob(plot)
# ind <- margtwoframes$layout$t[grepl("panel", margtwoframes$layout$name)]
# for (i in seq_along(ind)) {
#   margtwoframes$heights[[ind[i]]] = unit(fcsts[i], "null")
# }
# grid.newpage()
# grid.draw(margtwoframes)
```

>>>>

format:
  aft-pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
date: today
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
execute:
  echo: false
  eval: true

author:
  - name: Johannes B. Gruber
    affiliations:
      - name: University of Amsterdam
        department: Amsterdam School of Communication Research
        city: Amsterdam
        country: Netherlands
    email: J.B.Gruber@UVA.nl # corresponding author gets email
    orcid: 0000-0001-9177-1772
---

Word Count: **8978**

Keywords: *protest paradigm*, *Media framing*, *Social movements*, *News reporting*, *content analysis*

Contact:

Johannes B. Gruber\newline
J.B.Gruber@UVA.nl\newline
Nieuwe Achtergracht 166\newline
1018 WV Amsterdam, Netherlands\newline
+31 6 81963022

# Author biography

**Johannes B. Gruber** completed his PhD at the University of Glasgow before working as a postdoc-toral researcher at the Chair for Digital Democracy at the European New School of Digital Studies (ENS) and later the departments of Communication Science at Vrije Universiteit Amsterdam and the University of Amsterdam. His substantial research focuses on news media framing of protest and agenda setting in hybrid media systems. Beyond that, he is interested in text as data methods and tools.
